### 2017-08-02 pldi-2017/open-world-soundness-is-too-weak.tex
pldi: sketch trivial, open-world sound languages

\documentclass{article}
\input{texstyle.tex}
\begin{document}

%% TODO this is a good blog post

\begin{abstract}
  Reticulated's open-world soundness is a weak property because even a dynamically-typed,
   Python-like language is open-world sound.
\end{abstract}

\tableofcontents

\section{Python}

This is a nonsense typed language that compiles into Python.
It is nonsense for two reasons:
\begin{itemize}
\item The type-checker accepts every well-formed program
\item Evaluation can raise arbitrary blame errors (the programmer gets no hint as to what went wrong)
\item The environment $\Gamma$ is ignored.
\end{itemize}

\begin{mathpar}
  \inferrule*{
  }{
    \Gamma \vdash n \compilesto n : \tint
  }

  \inferrule*{
  }{
    \Gamma \vdash x \compilesto x : \tint
  }

  \inferrule*{
    \Gamma \vdash e_0 \compilesto e_0' : \tint
    \\\\
    \Gamma \vdash e_1 \compilesto e_1' : \tint
  }{
    \Gamma \vdash e_0 + e_1 \compilesto (\dyncheck{e_0'}{\kint}) +^\torigin (\dyncheck{e_1'}{\kint}) : \tint
  }

  \inferrule*{
    \Gamma \vdash e \compilesto e' : \tint
  }{
    \Gamma \vdash \tfun{f}{x}{T_1}{T_2}{e} \compilesto \ufun{f}{x}{e'} : \tint
  }

  \inferrule*{
    \Gamma \vdash e_0 \compilesto e_0' : \tint
    \\\\
    \Gamma \vdash e_1 \compilesto e_1' : \tint
  }{
    \Gamma \vdash (e_0~e_1) \compilesto ((\dyncheck{e_0'}{\kproc})~e_1')^\torigin : \tint
  }

  \inferrule*{
    \Gamma \vdash e \compilesto e' : \tint
  }{
    \Gamma \vdash \vref{e} \compilesto \vref{e'} : \tint
  }

  \inferrule*{
    \Gamma \vdash e \compilesto e' : \tint
  }{
    \Gamma \vdash !e \compilesto !(\dyncheck{e'}{\kref}) : \tint
  }

  \inferrule*{
    \Gamma \vdash e_0 \compilesto e_0' : \tint
    \\\\
    \Gamma \vdash e_1 \compilesto e_1' : \tint
  }{
    \Gamma \vdash e_0 \assign e_1 \compilesto (\dyncheck{e_0'}{\kref}) \assign e_1' : \tint
  }
\end{mathpar}

Here is another nonsense translation:

\begin{mathpar}
  \inferrule*{
    \fresh{l}
  }{
    \vdash e_s \compilesto \dyncheck{0}{\tagof{T}, l, \respath} : T
  }
\end{mathpar}

\section{Open World Soundness}

%% TODO should use "type soundness", the corollary

From POPL 2017

\textbf{Lemma B.13} (Open world soundness)
\begin{itemize}
\item[] \emph{If}
  \subitem $\Gamma \vdash e_s \compilesto e : T$
  \subitem $\vdash \context : \tagof{\Gamma} ; \tagof{T} \Rightarrow \emptyset; S$
\item[] \emph{then}
  \subitem $\emptyset; \emptyset \vdash \context[e] : S$
\item[] \emph{and either:}
\item[]
  \begin{itemize}
  \item $\config{\context[e], \emptyset, \emptyset} \stepsto \config{v, \sigma, \blameset}
         \mbox{ and }
         \emptyset; \Sigma \vdash v : S
         \mbox{ and }
         \Sigma \vdash \sigma$
  \item $\config{\context[e], \emptyset, \emptyset} \stepsto \blame{\labels}$
  \item $\config{\context[e], \emptyset, \emptyset} \stepsto \config{e', \sigma, \blameset}
         \mbox{ and }
         \config{e', \sigma, \blameset} \mbox{ is stuck}$
  \item $\config{\context[e], \emptyset, \emptyset}$ diverges
  \end{itemize}
\end{itemize}

Informally, if a Reticulated program $e_s$ is well-typed, then compiling the program,
 using it in any Python context, and running the whole program will yield one
 of four outcomes:
\begin{itemize}
\item evaluation produces a value with the same \emph{type-tag} as the source expression;
\item evaluation reduces to a blame error that points to a set of labels;
\item evaluation reduces to a segfault due to Python; or
\item evaluation diverges.
\end{itemize}

\subsection{Auxiliary Definitions}

From POPL 2017.
(This section should be more detailed.)

\begin{itemize}
\item[] $\compilesto$ translates a Reticulated program into a "safe" Python program
\item[] $\tagof{}$ gets the top-level constructor of a type
\item[] TODO context typing
\item[] TODO $\Sigma \vdash \sigma$
\item[] $\labels$ is a set of blame labels
\item[] $\stepsto$ is evaluation
\item[] TODO more???
\end{itemize}

\section{Python is Open-World Sound}

\section{Closing Remarks}

Dynamically typed \ldots won't get stuck \ldots blame is trivial \ldots labels always fresh \ldots still blame-subtyping

You need requirements!
How to rule out trivial things??!

\end{document}