### 2018-01-27 gf-pldi-2018/RESPONSE-PLDI-2018.txt
pldi: draft full response

Response Part 1: clarifications
================================================================================

From the "Paper summary" in Review #99A:

> This paper particularly considers five different embeddings, including the
> erasure embedding ... and locally-defensive embedding, which removes many
> unnecessary runtime checks from natural embedding

The locally-defensive embedding skips checks that are **necessary** for
type-sound execution, but unnecessary for the weaker type-tag-sound execution.
The evaluation then illustrates the tradeoff between soundness and performance.

Pycket is an orthogonal result
--------------------------------------------------------------------------------

From Review #99A:

> While the locally-defensive embedding in this paper improves the performance
> of the natural embedding by about a magnitude, the recent work by Pycket has
> already achieved this.

The Pycket work demonstrates that replacing a conventional JIT with a tracing
JIT can significantly reduce the overhead of natural-embedding type soundness.

Our work demonstrates that replacing type soundness with type-tag soundness
can lead to significant performance improvements.

These are separate results.

Future work: combine these two independent pieces of work, which will require
close collaboration with the Pycket team.

Locally-defensive is a principled transient
--------------------------------------------------------------------------------

From Review #99A:

> Maybe transient typed Racket would be a good direction to make gradual typing
> (migratory typing) for Racket more feasible?

From Review #99C:

> There is perhaps another contribution: the idea of "locally-defensive"
> embedding.  Frustratingly, I can't figure out if this is an original
> contribution of the paper, or if it's taken from somewhere else.

The locally-defensive embedding is a mathematical characterization of
transient semantics, and our locally-defensive Racket implements it.

In contrast, the POPL 2017 paper by Vitousek et al. defines transient by its
implementation.  That paper does not explain the properties that an
implementation needs to satisfy in order to be called "transient" --- so it is
unclear, for example, how to give a transient semantics for the gradually-typed
lambda calculus.  Our work fills this gap.

POPL 2017 does not systematically compare type soundness and tag soundness
--------------------------------------------------------------------------------

From Review #99A:

> Also, the transient strategy of ensuring type soundness exhibits much more
> competitive performance.

From Review #99B:

> The main result is a confirmation that the transient semantics of Vitousek
> (POPL 2017) is indeed better-performing than the usual semantics with eager
> casts. This is hardly surprising given the experiments in that paper.

Review #99B sells our results short:

1. Our paper provides a mathematical comparison between type soundness and tag
   soundness (Vitousek et al. conflate the two; see their Corollary 5.5.1).
2. Our paper provides the first controlled comparison of type soundness and
   type-tag soundness.
3. Vitousek et al. do not report the performance of programs that mix typed
   and untyped code.

Our paper is aiming to be scientific, not surprising.

The idea of a spectrum has NOT been previously explored
--------------------------------------------------------------------------------

From Review #99B:

> The other claimed contribution is the idea that there is in fact a spectrum
> of soundness and performance: but this has been previously explored in the
> work on like types, at least.

Review #99B mischaracterizes like types. Like types describe a novel type
system with a conventional notion of type soundness.

Our paper is the first to show how _one language_ can have _multiple semantics_
each corresponding to a _different kind of soundness_.

Sound gradual typing is a worthwhile goal
--------------------------------------------------------------------------------

From Review #99C:

> There is a big literature on gradual typing, but only one form appears to have
> had practical impact.  That's the one that is here called the "erasure
> embedding".  The obvious question is: are the other approaches worth pursuing
> at all?

Industry chose erasure to avoid the performance question.  Experience, and our
theorem (L_E term safety), show that this choice lets programs produce wrong
answers due to uncaught violations of the static types.  This has two negative
consequences: users need to debug these subtle errors, and users cannot use
types to reason about correctness.

We are pursing sound gradual typing because we believe it is a useful
improvement over optional typing.  It addresses the above issues.

In the same way, the implementors of Scala and Haskell believe that static
typing is a useful improvement over the dynamic typing in, e.g., Ruby and
Python, despite Ruby and Python having a larger practical impact.

Object-oriented features are future work
--------------------------------------------------------------------------------

From Review #99E:

> The paper does not apply these approaches to the object-oriented features of
> Racket ... Do you anticipate the locally-defensive approach will help in that
> setting?

Yes.  Based on the theory, the difference is between an expensive object proxy
and a near-constant-time check.  We did not implement tag soundness for objects
because (1) implementing tag soundness for functions was already a significant
effort and (2) type-sound objects for Typed Racket required a dissertation to
design and implement.  It's on our agenda, but was not our main focus for this
paper.

Response Part 2: answers
================================================================================

Review #99A
--------------------------------------------------------------------------------

> how hard it is to implement the idea of the forgetful and co-natural
> embeddings on top of locally-defensive embedding.

The locally-defensive embedding builds on the ideas of the forgetful and
co-natural embeddings.  The first paragraph of section 3.8.1 tries to explain
the connnection.

> How hard is it to integrate co-natural and forgetful embeddings into the
> current prototype for locally-defensive embedding?

Implementing the co-natural and forgetful embeddings for Racket would be a
significant undertaking.  The co-natural embedding requires new kinds of
proxies, for which one would need to modify the Racket runtime.  The forgetful
embedding requires significant changes to Racket's contract library --- one
would need to modify function contracts so their domain and codomain could be
replaced at runtime.

> Line 153, why parentheses around dyn $\tau$ e?

To clearly separate the code from the prose.  There should be parentheses around
stat $\tau$ e on this line as well.

> Line 360, v should $v \colon \tau$.

Yes, thank you.

> Line 362, could you please provide an example expression such that
> e is well typed and reducing it is divergent?

No, we cannot provide an example.
Instead we could prove a theorem that L_S is strongly normalizing and remove the
clause on line 362; however, this is orthogonal to the focus of the paper.

> Figure 4, what is the meaning of $A \notin e'$? Did you mean A is not 
> a TagError and BoundaryError in the first occurrence and A is not a 
> BoundaryError in the second occurrence? In particular, what does 
> $e'$ bind to? Please clarify.

This figure has a typo: it should say $A \not\in e'$.  By this we mean `A` is
not a `TagError` and not a `BoundaryError`; specifically, the answer value `A`
is a member of the set `e` defined by the grammar at the top of figure.
(And yes, the grammar uses the symbol `e` for both the set of expressions and
for a member of the set.)

Review #99B
--------------------------------------------------------------------------------

Review #99C
--------------------------------------------------------------------------------

> The other frustrating thing about the locally-defensive embedding is that I
> never really understood it.

The goal of the locally-defensive embedding is to guarantee type-tag soundness
(section 3.8.3).  The locally-defensive embedding implements this goal by
rewriting statically-typed expressions to check the type-tag of every incoming
value.  For the language in figure 9, "incoming" values come from `dyn`,
function applications, and calls to `fst` and `snd`.

> The performance measurements are indicative but not conclusive: it's always
> hard to know what to conclude about absolute speedups when so much depends on
> the constant factors, and may be dramatically affected by decent compilation.

To clarify: we report relative performance, we use the same compiler and JIT
for all programs and soundnesses, and these programs are running computations
of at least a few seconds in the fully-untyped configuration.  (Notably, `fsm`
and `zombie` run short computations because some of their typed/untyped
configurations run exponentially slower.)

> For example, perhaps a simple static analysis of a natural embedding could
> remove many checks.

We, and others, have invested significant effort in improving Typed Racket
since the POPL 2016 paper "Is Sound Gradual Typing Dead".  In fact, the data
for Typed Racket in our submission is a significant improvement over the data
in POPL 2016.  We have been unable, however, to find a static analysis that
leads to an order-of-magnitude improvement.

> given the maturity of the field I'd appreciate the authors' perspective on
> the potential impact of their work.

Our results suggest that type-tag soundness is worth adding to TypeScript so
that programmers get a (weak) soundness guarantee without much change in
performance.  Experience with tag soundness may lead to a demand for type
soundness.

> Can we assume that proofs exist?  Are they typeset somewhere?

Proofs are typeset, and the Redex models in our submission have QuickCheck-style
tests to validate the soundness theorems.

> I'd have thought that a helpful application of your framework would be to
> identify hot-spots: values that move to and fro a lot, or "if you just make
> this bit of code statically checked, then your whole program would run
> faster".   A kind of gradual profiler.  You have the machinery in hand...
> and it'd be useful

We are working on it.

> Lines 245-250, The last para before 3.1, about the meta function that lifts a
> notion of reduction, was pretty obscure.  I read it several times, but never
> really "got it".

We hope figure 1 was easier to read.  For reference, the citation for this
sentence is meant to direct readers to chapter 5, section 1 of _Semantics
Engineering with PLT Redex_.

> Line 256 "..if it contains no free variables".  No: if al the free variables
> of e are bound by \Gamma.

Right, if `e` contains a variable that is not bound by `\Gamma` then the rules
in figure 2 cannot be used to prove `\Gamma \vdash_D e`

> line 265 "...representative of subset types that do not have a matching low
> level type tag".  I have no idea what this means.

We need to make this sentence clearer.

The idea is that most languages with runtime type-tags, such as OCaml, normally
have a tag to say "these bits represent an integer" and no tag to say "these
bits represent a natural number".  Therefore, the runtime notion of type-tags
in such a language needs to be generalized to support a gradual type system
that distinguishes Int and Nat types.

> line 380 "...definition of LM  does not include a semantics".  What does it
> mean to "include a semantics".  By "a semantics" I think you mean a definition
> of R and R'.

By "include a semantics", we mean "to define a notion of reduction R and a
reduction relation `e -->_R^* A`.  In contrast, figures 2 3 5 6 7 and 8 all
"include a semantics".

> (Incidentally, why use R and R' for static and dynamic rather than R^S and R^D?
> I kept stumbling over which was which.)

This is a good suggestion.

> Line 284. I have not clue what the bow-tie thing is doing.  It did not seem
> to be used subsequently.

Figures 6 7 and 8 use the bow-tie to combine two notions of reduction (one for
statically-typed expressions, one for dynamically-typed expressions) into two
reduction relations (one for static, one for dynamic) that are defined for
expressions that mix statically-typed and dynamically-typed code.

The bow-tie helps us prove that a "static" reduction step cannot end in a
`TagError`.

> 3.6 (co-natural) and 3.7 (forgetful).  co-natural affects semantics (line 655)
> but I believe that forgetful does not.  Correct?  Please say so.

The semantics for co-natural is defined in figure 7.  The semantics for
forgetful is defined in figure 8.  These are different semantics.

> The type-safety theorems have an identical form in each case (same in 3.5) so
> it'd be helpful to have a formal statement of how they differ!  Perhaps, if
> the natural embedding converges to a value then the co-natural one does too,
> and to the same value?

This is a good suggestion.

We have to be careful, though, about what it means for a value to be the same.
The expression `(dyn (Nat x Nat) (1, 1))` reduces to `(1, 1)` in the natural
embedding and to `(mon (Nat x Nat) (1, 1))` in the co-natural embedding.

Sub-Review #99C
--------------------------------------------------------------------------------

> The authors claim at least 5 different soundness conditions (line 1284), do
> these correspond to the 5 languages (LE, LN, LC, LF, LK)?

Yes.

> From my understanding 3 of them have the same soundness condition (based on
> ... line 747)

Line 747 is a typo, it should say "a similar notion of soundness".

> Perhaps a table that captures the differences between each language would be
> nice.

Yes.

> Line 70: I'm unsure if they mean languages like TypeScript ... If so, calling
> them gradual typing rather than optional typing could be misleading.

We should clarify this.

> Line 648: Should Theorem (L_C type safety) read type soundness?

Yes.

> Line 703 till end of paragraph: I feel this could have more
> explanation; the result that you can forget all but one monitor isn't
> immediately obvious.

Yes, we should clarify.

> Line 1003: Are they supporting parametric polymorphism? I'm not convinced
> that it is possible to enforce parametricity with the monitor-free,
> tagging semantics.

No, we are not claiming to support parametricity.  That is a separate and very
difficult research question (Devriese et al. POPL 2018).

> Line 1023: 'By contrast, we view the co-natural and forgetful embeddings as
> theoretical artifacts.' I have 2 possible queries ...

Good point, we should re-word.

Review #99D
--------------------------------------------------------------------------------

> no input from actual developers is ever collected (user studies, observations
> via check-ins, etc.) so it is quite difficult to hypothesize both the amount
> of labor that goes into annotations and the demands in terms of acceptable
> overhead.

The annotation effort is orthogonal.  Our semantics re-use Typed Racket's type
checker.  The annotation burden of Typed Racket has been reported elsewhere,
for example, in chapter 11 of Tobin-Hochstadt's dissertation.

Users' demands in terms of acceptable overhead is not a major concern at this
point.  We are focused on improving the performance of sound gradual typing;
once we run out of improvements, then we will ask users whether the current
performance meets their needs.

> Lastly, the level of difficulty that's involved in pipointing a
> type violation is also hard to measure -- very much depends on the size and
> structure of the program

True, but one can easily compare the quality of error messages.

> I like the notion of D-deliverable configurations from [27], however, no
> guidance is given for selecting D in practice even for the small and simple
> benchmarks that you show

Section 5.2 says "The premise of the D-deliverable measure is that programmers
have a fixed performance requirement".  In other words, we assume that a
programmer (or their boss) has a requirement, e.g., we can tolerate at most
a 4x slowdown without losing customers.

Review #99E
--------------------------------------------------------------------------------

> Figure 1. What does the notation "A not in e" mean in the side condition of the
> evaluation rule for E[e] evaluating to A?

See our response to Review #99A above.
In short, the notation means that `A` is not a `TagError` or `BoundaryError`.

> Figure 4.  How does the dynamic language know the type of the context into
> which e is being embedded?

The type is part of the syntax of embedded terms.  In the model, embedded terms
have the form `(dyn \tau e)` or `(stat \tau e)`.  In the implementation, the
compiler uses type annotations and inferred types to handle the embedding.

> Figure 4.  Explain the notation-in-box "inherits" judgement: notation.

By "inherits", we mean "recursively extends".  William Cook's SBLP slides
give an overview:
  <http://www.cs.utexas.edu/~wcook/Drafts/2009/2009-SBLP-Inheritance.pdf>

> Line 461.  What are category I and category II disasters?

Sorry for the jargon --- we borrowed these terms from the Saffir-Simpson
hurricane wind scale.  We used them to say "bad" and "worse".  In the future
we will use plain English.

> Line 463.  Do you mean for the second type in the pair to be Nat when the
> professor said it was the first element that was non-negative ... ?

Yes, we made a mistake.

> Line 685.  Could you cache the result of a successful check so the cost for all
> but the first check would be negligible?

Yes, this is a promising idea.

But as Findler et al. IFL 2007 show for lazy contracts, adding a cache may
affect performance in a complicated way.

> Line 705.  This sentence doesn't make sense.

Correct, this sentence is missing the phrase "combines the".

> Line 904.  Can you give an example of a well-typed term that isn't well-tagged
> with the corresponding tag?

`(fst (1, 1))`