### 2018-05-04 gf-icfp-2018/RESPONSE-ICFP-2018.txt
revised icfp response

ICFP '18 Paper #21 Author Response
===========================================================================

Review #21B mischaracterizes our results in at least three ways.

FIRST, the review misrepresents our theoretical results as variations of
known theorems.

This claim is _completely wrong_ (with an exception). Our theorems expose
the looseness of previous claims in the literature concerning "soundness"
in the area of gradual typing. Our meta-theorems (and their logical
implications) explain that these claims are incomparable and must be read
very carefully. Technically our meta-theorems show that soundness results
can be stated only in terms of properties preserved through run-time
configurations. As a result, the implications for the programmer differ
dramatically. For details, see below.

SECOND, the review #21B seems to imply that our results might be preempted
by two papers (one of which is still unavailable in public):

> _A Framework for Object-Oriented Gradual Typing_.
>  Chung, Li, Zappa Nardelli, Vitek.
>  ECCOP 2018
>
> _Efficient Gradual Typing_.
> Kuhlenschmidt, Almahallawi, Siek.
> https://arxiv.org/abs/1802.06375

Felleisen is the PI on a large NSF grant that also supports these efforts.
The papers were presented at the August 2017 PI day at Northeastern. When
we submitted our paper, Chung et al.'s work was NOT ACCEPTED and NOT
AVAILABLE in public. [[ We confirmed this statement with the authors who
sit across the hallway from us. ]]  Now that its status has changed we can
easily compare the two research results and will do so in the paper. For
a detailed comparison, see below.

We do not consider Kuhlenschmidt et al.'s extremely preliminary work
related and we do not accept that it contradicts our result. In brief,
Kuhlenschmidt et al. are working on an alternative approach to the
compilation of NEWLY DESIGNED gradually typed languages. Their work has
confirmed that on a very small number of micro-benchmarks, the compiler can
eliminate a large fraction of overhead due to the TRANSIENT semantics of
gradual semantics.

Kuhlenschmidt et al.'s work does not include the apples-to-apples
comparison of work, with full-fledged implementations of all three major
semantics. Similarly, while our three implementations can deal with a dozen
non-trivial benchmarks (programs used on a daily basis), Kuhlenschmidt et
al.'s compiler cannot deal with anything realistic so far. Of course, we do
look forward to the day when their work supersedes the current state of the
art with respect to performance.  We wish to use gradually typed languages
on a daily basis.

THIRD, the review claims that our paper does not support our statements
about performance. Specifically, the review claims

> p19. "Sound static types can eliminate the need to dispatch, thus the
> natural embedding's performance can exceed that of the erasure
> embedding"
> This paper does not provide empirical results regarding this claim.

The reviewer seems to have overlooked TABLE ____ttt___ on PAGE ___ppp___.
This table clearly demonstrate that the performance of NATURAL dominates
the performance of LOCALLY-DEFENSIVE on programs whose main modules are
fully typed (though they may use untyped run-time libraries).

Response to Review #21A
===========================================================================

> A weak part of the paper---which is easy to fix---is that there is no
> discussion of mixed-semantics languages.  As such, Dart is badly characterized
> in this paper as an optional/erasure style language.

Thank you for reminding us of Dart's mixed approach. We will include a
brief discussion and draw attention to its approach, also as an area for
future work.

> - Please explain why errors do not propagate through E^\bullet contexts

Errors do propagate through, because E^\bullet is a subset of E.

> - In Theorem 2.0, 2nd reduction alternative, I don't understand why the right
>   hand side cannot be simply: E[dyn t' E* [TagErr]], considering that e' steps
>   to TarErr and that the reduction relation admits reduction under E*.
>   (same question for Theorem 2.6)

You are correct. Thanks!

> - The data of Figure 12 could be plotted (again, there is space)

Yes! This table is too easy to miss.

Response to Review #21C
===========================================================================

> I am a little bit surprised at the fact that static typing does not
> contribute the runtime performance at all.  I would like to see the
> runtime performance on a JIT compiler or a compiler equipped with
> type-conscious optimization.

This is a correct observation. The Indiana Typed Racket team is revising
the Pycket compiler for Racket to address this issue. We are looking
forward to reproducing our experimental work on top of that new
backend. But this area remains future work for sound gradually typed
approaches.

Detailed Response to Review #21B
===========================================================================

> + The paper does a thorough and controlled performance experiment that
>   confirms that the results of Vitousek et al. (2017) on Reticulated
>   Python carry over to the setting of Typed Racket. That is, the
>   transient semantics avoids the catastrophic performance problems
>   that occur in naive implementations of the natural approach, but
>   transient incurs overhead in statically typed code.

Vitousek et al. (2017) does not directly compare transient to natural.

Remarks like this are exactly why we prepared this paper. It is time to stop
polluting the scientific literature with general claims about different
languages with different type systems and different guarantees.

> * The semantic framework is a straightforward adaptation of
>   the Matthews-Findler multi-language semantics.

The first sentence of Section 2.2 and first paragraph of Section 6
(the conclusion) acknowledge Matthews & Findler. If this is not enough,
we can add more.

> * Theorem 2.0 is a just a combination of the usual Type Safety theorem
>   for gradually typed languages (Siek and Taha 2006) and the Blame
>   Theorem (Tobin-Hochstadt and Felleisen 2006, Wadler and Findler
>   2009). The paper does not discuss these relationships.

The paper does not claim that these theorems are novel. The important
contribution is expressing the different approaches in a framework that
draws attention to their meaningful distinctions.

That said, the difference between Theorem 2.0 and prior work is that it
explicitly connects the surface language (without proxy/monitor values) to the
core language (with proxy values). This is an important distinction for
language designers: even if the surface language does not expose proxies,
users need to understand proxies and the evaluation type system to understand
the language's soundness guarantee.

> * Theorems 2.1, 2.3, and 2.4 are all the usual type safety theorem for
>   dynamically typed languages.

Again, the paper does not claim these theorems are novel but it is definitely
unusual that these theorems include 2 typing judgments.

These theorems also emphasize that soundness for a pair of languages comes
as a pair of theorems. (Non-experts should be wary of papers that advertise
a "Theorem 2.5"-style result without discussing partially-typed programs.)

> * Theorem 2.5 is just type safety for a statically typed program, and
>   is a corollary of the equivalence between gradual typing and static
>   typing on fully annotated programs (Siek et al. 2015).  The paper
>   does not discuss this close relationship.

Good point, we will discuss the relationship.

> * Theorem 2.6 is an adaptation of Corollary 5.5.1 (Type soundness) of
>   Vitousek et al. (2017).  The paper does not discuss this close
>   relationship.

Theorem 2.6 is **not** an adaptation of Vitousek et al.'s "type soundness".
Vitousek et al.'s Corollary 5.5.1 is about a single language.

Nor is our Theorem 2.6 an adaptation of Vitousek et al.'s Theorem 5.5
(Open-World Soundness). There are two important differences:

- Vitousek et al. do not define a type system for the surface language;
  it is impossible to reason about programs independent of their core-language
  translation (via Vitousek's $\rightsquigarrow$).

- Vitousek et al.'s "stuck terms" are defined in terms of the core language;
  it is not clear from the theorem that typed programs cannot get stuck without
  studying their translation into the core language. In contrast, our theorem
  shows that only dynamically-typed expressions can reduce to a TagError.

We will add a clarification to the paper.

> * Theorem 2.7 should be another example of the usual type safety
>   theorem for a dynamically typed language, but it looks like there
>   is a typo in saying |-_{LD} v : \lfloor \tau \rfloor
>   because there is no \tau to refer to here.

Thank you, that is definitely a typo.

> p1. The abstract is much too short.

The abstract says just what we choose it to say --- nothing more.

If you believe something is missing from the abstract, we will consider
including it.

> p1. "But, it may impose a huge run-time cost"
> It would be appropriate to cite recent work by Kuhlenschmidt et
> al. (see above) that shows that the natural semantics does not have to
> impose a huge run-time cost.

We disagree, it would not be appropriate to cite Kuhlenschidt et al. here.
The sentence you quote is trying to communicate the historical fact that
the performance of the natural embedding led researchers and implementors
to consider other alternatives.

Kuhlendschmidt et al.'s work does not change the past.

> p1. In the discussion of "type soundness", this paper should discuss
> the Blame Theorem, which provides the same guarantees that are provided
> by the theorems in this paper.

All the blame theorems we know (see below) are guarantees about casts in an
intermediate language. Our type soundness theorems are guarantees about the
results of evaluating a surface-language term. These are different guarantees.

"The Blame Theorem"s (all of which are slightly different):

- Wadler and Findler (2009)
- Ahmed et al. (2011)
- Vitousek et al. (2017)

> p9. "Second, monitoring adds prohibitive run-time cost."
> That is only true of naive implementations. Again, see
> Kuhlenschmidt et al.

Kuhlenschmidt et al's work is not relevant here --- the sentence you quote is
summarizing the motivation for transient given in Vitousek et al. (2017).

Please let us know if you believe we have mis-represented Vitousek et al. (2017)

> p4. It is confusing to use boundary errors for both cast errors and
> for errors such as div-by-zero.

We do this to show how soundness for a pair of languages is a natural
generalization of soundness for a single language that includes primitive
operations (such as division) in its trusted code base.

We will clarify in the paper.

> p9. "near-constant time"
> Why only "near" and not just "constant time"?

One reason is for union types. The check for `(U String Integer)` must perform
at most two constructor checks at runtime.

Another reason, from Vitousek et al. (2017), is for structural object types.
Checking that an object has the expected members requires one check for each
member in the structural type.

> p10.
> Why do you perform checking of parameters in the reduction rules
> but insert all the other checks during compilation? This seems
> rather inconsistent and differs, as you later point out,
> from the implementations.

Footnote 3 (line 537) explains.

> p16.
> "This table demonstrates that for pathological examples..."
> Those examples are not pathological.

Suggestions welcome. We just want a word to describe "examples with worst-case
performance".

> p17. "the implicit conjectures of the literature"
> You seem to be overlooking the performance experiments
> reported on by Vitousek et al. (2017).

Vitousek et al. (2017) do not measure the performance of programs that mix
typed and untyped code.

> p19. "The natural embedding incurs three significantly larger kinds of
> costs." You're forgetting the fourth: the natural embedding also has
> the space-efficiency problem pointed out by Herman et al. (2006),
> which also induces problems with time efficiency.

We did not forget, but you are right we need to be clearer that the cost
of **allocation** is a space+time cost and is unbounded in principle.

We will clarify.

> p19. "Sound static types can eliminate the need to dispatch, thus the
> natural embedding's performance can exceed that of the erasure
> embedding"
>
> This paper does not provide empirical results regarding this claim.

See Figure 12.